---
title: "MPC"
author: "NCompton"
date: "28/06/2021"
output: html_document
---

knit at end

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

install packages and set libraries
```{r, echo = FALSE}
#install.packages("odbc")
#install.packages("DBI")
#install.packages("dplyr")
#install.packages("fitdistrplus")
#install.packages("ggplot2")
#install.packages("qicharts2")
#install.packages("psych")
#install.packages("distributions3")
library(psych)
library(qicharts2)
library(ggplot2)
library(fitdistrplus)
library(dplyr)
library(odbc)
library(distributions3)
```

Define functions
is the z test paired? 
```{r}
# function to check the distribution of the data input
check_distribution <- function(data){
  descdist(data, discrete = FALSE)
  normal_dist <- fitdist(data, "norm")
  plot(normal_dist)
}




# function to return 2-sample z statistic and p value
z.test <- function(x, y){
  # get the sample mean, sd and n
  mu_1 = mean(x)
  mu_2 = mean(y)
  sigma_1 = sd(x)^2
  sigma_2 = sd(y)^2
  n_1 = length(x)
  n_2 = length(y)
  
  # calculate 2 tailed z stat
  z_stat = (mu_1 - mu_2) / (sqrt ((sigma_1/n_1) + (sigma_2/n_2)))
  
  # calculate p value
  Z <- Normal(0,1) # make a standard normal r.v
  p = 2 * cdf(Z, abs(z_stat))
  return(c(z_stat, p))
}

z.paired <- function(x, y, d0){
  mean_diff <- mean(x-y)
  sigma_diff <- sqrt(var(x)+var(y)-2*cov(x,y))
  n <- length(x)
  z<-sqrt(n)*((mean_diff-d0)/sigma_diff)
  p_value=2*pnorm(-abs(z))
  
    # qnorm gives the alpha quantile of the standard normal  distribution
  # reject H0 that difference of means = d0 if z < z(alpha/2) or
  # z > z(1-alpha/2)
  alpha = 0.05
  alpha2 = 0.01
  
  
  if(z < qnorm(alpha/2)) {print("Null hypothesis can be rejected. Evidence the difference in means is not equal to", d0)}
  if (z > qnorm(1 - (alpha/2))) {
  print("Null hypothesis can be rejected. Evidence the difference in means is not equal to", d0)}
  else {  print("Null hypothesis can not be rejected, evidence the difference in means is equal to", d0)}
  
  return(c(z, p_value))
}
```


Set up the database connection
```{r}
con <- DBI::dbConnect(odbc::odbc(), 
                      Driver = "SQL Server", 
                      Server = "IT049561\\SQLEXPRESS", 
                      Database = "MPC", 
                      Trusted_Connection = "True")
```


Get a dataframe for each beam energy
```{sql connection=con, output.var = "df6x"}
SELECT *
FROM "6x"
```
```{sql connection=con, output.var = "df10x"}
SELECT *
FROM "10x"

```
```{sql connection=con, output.var = "df10fff"}
SELECT *
FROM "10fff"

```

Get the dataframes of the recent data - from 1st April 2021 when we measured the water phantom
```{sql connection=con, output.var=recent6x}
SELECt *
FROM "6x"
WHERE QA_Date >= '2021-04-01'
```

```{sql connection=con, output.var=recent10x}
SELECt *
FROM "10x"
WHERE QA_Date >= '2021-04-01'
```

```{sql connection=con, output.var=recent10fff}
SELECt *
FROM "10fff"
WHERE QA_Date >= '2021-04-01'
```

Descriptive statistics with psych
```{r, echo = FALSE}
describeBy(df6x$MPC_Distance)

describeBy(df6x$Calibrated_Distance)

describeBy(df6x$Uniformity)

describeBy(df6x$Output_Change)
```

check distributions are normal - irrelevant since we have large sample sizes (over 30)


Comparing the symmetry and flatness before and after the calibration applied; and in inline and crossline directions
We use paired z tests
Z test null hypothesis: the means of the two groups are equal
```{r}
#z.test(df6x$MPC_Distance, df6x$Calibrated_Distance)
#z.test(df6x$SymmInAT, df6x$SymmCrossAT)
#z.test(df6x$SymmCrossBT, df6x$SymmCrossAT)
# p values greater than 1?

#z.paired(df6x$SymmInAT, df6x$SymmInBT, 0)
#z.paired(df6x$SymmCrossAT, df6x$SymmCrossBT, 0)
#z.paired(recent6x$SymmInAT, recent6x$SymmCrossAT, 0)

#z.paired(recent6x$FlatInAT, recent6x$FlatCrossAT, 0)
#z.paired(recent6x$FlatInBT, recent6x$FlatInAT, 0)

boxplot(df6x$SymmInAT, df6x$SymmCrossAT)
z.paired(df6x$SymmInAT, df6x$SymmCrossAT, 0)
pnorm(-1.96)
```

Taking a random sample from the data
```{r}
set.seed(2) # set seed makes the analysis reproducible
rand_df <- df6x[sample(nrow(df6x), size=300), ]
z.paired(rand_df$SymmInAT, rand_df$SymmCrossAT, 0)
```


beam centre shift
```{r}
wilcox.test(df6x$MPC_Distance, df6x$Calibrated_Distance)
wilcox.test(df10x$MPC_Distance, df10x$Calibrated_Distance)
wilcox.test(df10fff$MPC_Distance, df10fff$Calibrated_Distance)

wilcox.test(recent6x$MPC_Distance, recent6x$Calibrated_Distance)
wilcox.test(recent10x$MPC_Distance, recent10x$Calibrated_Distance)
wilcox.test(recent10fff$MPC_Distance, recent10fff$Calibrated_Distance)
```

Thresholds for passing the MPC tests are
centre shift = 0.5
OutputChange = 2
Uniformity = 2
Create a conditional column for if it passed/failed on my algorithm and on mpc
will this give sensitivity and specificity ?
or it will give another metric for comparative stats
there's no mpc that fail and loads of calibrated that do
should there be a cut off?
```{r}
mpc_6x$mpc_distance_pass <- "fail"
mpc_6x$mpc_distance_pass[mpc_6x$MPC_Distance<0.5 & mpc_6x$MPC_Distance> -0.5] <- "pass"
```

Add column which will be a pass/fail for the mpc uniformity and one for if any of the 4 symm/flat fails
or for how sensitive the symm is to uniformity and then how sensitive flatness is to uniformity
Then we can do sensitivty and specifity
```{r}
df6x$uniformity_test <- "fail"
df6x$uniformity_test[df6x$Uniformity<2 & df6x$Uniformity>-2] <- "pass"

df6x$symm_test <- "fail"
df6x$symm_test[df6x$SymmInAT<2 & df6x$SymmInAT>-2 & df6x$SymmCrossAT<2 & df6x$SymmCrossAT>-2] <- "pass"


recent6x$uniformity_test <- "fail"
recent6x$uniformity_test[recent6x$Uniformity<2 & recent6x$Uniformity>-2] <- "pass"

recent6x$symm_test <- "fail"
recent6x$symm_test[recent6x$SymmInAT<2 & recent6x$SymmInAT>-2 & recent6x$SymmCrossAT<2 & recent6x$SymmCrossAT>-2] <- "pass"

false_pos <- sum(recent6x$uniformity_test == "pass" & recent6x$symm_test == "fail")
true_pos <- sum(recent6x$uniformity_test == "pass" & recent6x$symm_test == "pass")
false_neg <- sum(recent6x$uniformity_test == "fail" & recent6x$symm_test == "pass")
true_neg <- sum(recent6x$uniformity_test == "fail" & recent6x$symm_test == "fail")

sens <- true_pos / (true_pos + false_neg)
print(sens)

spec <- true_neg / (true_neg + false_pos)
print(spec)

false_pos <- sum(df6x$uniformity_test == "pass" & df6x$symm_test == "fail")
true_pos <- sum(df6x$uniformity_test == "pass" & df6x$symm_test == "pass")
false_neg <- sum(df6x$uniformity_test == "fail" & df6x$symm_test == "pass")
true_neg <- sum(df6x$uniformity_test == "fail" & df6x$symm_test == "fail")

sens <- true_pos / (true_pos + false_neg)
print(sens)

spec <- true_neg / (true_neg + false_pos)
print(spec)
```
```{r}
df6x$flat_test <- "fail"
df6x$flat_test[df6x$FlatInAT<104 & df6x$FlatCrossAT<104] <- "pass"

recent6x$flat_test <- "fail"
recent6x$flat_test[recent6x$FlatInAT<104 & recent6x$FlatCrossAT<104] <- "pass"

false_pos <- sum(recent6x$uniformity_test == "pass" & recent6x$flat_test == "fail")
true_pos <- sum(recent6x$uniformity_test == "pass" & recent6x$flat_test == "pass")
false_neg <- sum(recent6x$uniformity_test == "fail" & recent6x$flat_test == "pass")
true_neg <- sum(recent6x$uniformity_test == "fail" & recent6x$flat_test == "fail")

sens <- true_pos / (true_pos + false_neg)
print(sens)

spec <- true_neg / (true_neg + false_pos)
print(spec)

false_pos <- sum(df6x$uniformity_test == "pass" & df6x$flat_test == "fail")
true_pos <- sum(df6x$uniformity_test == "pass" & df6x$flat_test == "pass")
false_neg <- sum(df6x$uniformity_test == "fail" & df6x$flat_test == "pass")
true_neg <- sum(df6x$uniformity_test == "fail" & df6x$flat_test == "fail")

sens <- true_pos / (true_pos + false_neg)
print(sens)

spec <- true_neg / (true_neg + false_pos)
print(spec)
```
sens/spec of beam centre shift
the MPC distance is always within 0.5mm threshold
```{r}
df6x$centre_mpc_test <- "fail"
df6x$centre_mpc_test[df6x$MPC_Distance<0.5 & df6x$MPC_Distance>-0.5] <- "pass"

df6x$centre_cal_test <- "fail"
df6x$centre_cal_test[df6x$Calibrated_Distance<0.5 & df6x$Calibrated_Distance>-0.5] <- "pass"

recent6x$centre_mpc_test <- "fail"
recent6x$centre_mpc_test[recent6x$MPC_Distance<0.5 & recent6x$MPC_Distance>-0.5] <- "pass"

recent6x$centre_cal_test <- "fail"
recent6x$centre_cal_test[recent6x$Calibrated_Distance<0.5 & recent6x$Calibrated_Distance>-0.5] <- "pass"

false_pos <- sum(recent6x$centre_mpc_test == "pass" & recent6x$centre_cal_test == "fail")
true_pos <- sum(recent6x$centre_mpc_test == "pass" & recent6x$centre_cal_test == "pass")
false_neg <- sum(recent6x$centre_mpc_test == "fail" & recent6x$centre_cal_test == "pass")
true_neg <- sum(recent6x$centre_mpc_test == "fail" & recent6x$centre_cal_test == "fail")

sens <- true_pos / (true_pos + false_neg)
print(sens)

spec <- true_neg / (true_neg + false_pos)
print(spec)

false_pos <- sum(df6x$centre_mpc_test == "pass" & df6x$centre_cal_test == "fail")
true_pos <- sum(df6x$centre_mpc_test == "pass" & df6x$centre_cal_test == "pass")
false_neg <- sum(df6x$centre_mpc_test == "fail" & df6x$centre_cal_test == "pass")
true_neg <- sum(df6x$centre_mpc_test == "fail" & df6x$centre_cal_test == "fail")

sens <- true_pos / (true_pos + false_neg)
print(sens)

spec <- true_neg / (true_neg + false_pos)
print(spec)
```

plot the beam output change over time
```{r}
ggplot(df6x, aes(QA_Date, Output_Change)) +
  geom_point() +
  xlab("QA Date") +
  ylab("Output Change") +
  ggtitle("Output Changes against Time")

```
polynomial regression model

```{r}
df6x$index <- 1:length(df6x$Output_Change)

fit3 <- lm(Output_Change ~ poly(index , 56, raw=TRUE), df6x)  
fit4 <- lm(Output_Change ~ poly(index , 14, raw=TRUE), df6x)  ## quartic
summary(fit3)

plot(df6x$Output_Change, main="plot", xaxt="n")
tck <- c(1, 50, 100, 150)
axis(1, tck, labels=FALSE)
mtext(df6x$QA_Date[tck], 1, 1, at=tck)
lines(fit3$fitted.values, col=3, lwd=2)
lines(fit4$fitted.values, col=2, lwd=2)
legend("topleft", c("cubic", "quartic"), lwd=2, col=3:2)
```

Polynomial regression on this year

```{r}
recent6x$index <- 1:length(recent6x$Output_Change)

fit <- lm(Output_Change ~ poly(index , 56, raw=TRUE), recent6x)  
fit4 <- lm(Output_Change ~ poly(index , 14, raw=TRUE), recent6x)  ## quartic
summary(fit3)

plot(recent6x$Output_Change, main="plot", xaxt="n")
tck <- c(1, 50, 100, 150)
axis(1, tck, labels=FALSE)
mtext(recent6x$QA_Date[tck], 1, 1, at=tck)
lines(fit$fitted.values, col=3, lwd=2)
lines(fit4$fitted.values, col=2, lwd=2)
legend("topleft", c("cubic", "quartic"), lwd=2, col=3:2)
```


SPC charts for the beam centre shift over time;
the flatness and symmetry over time
```{r}

qic(df10x$MPC_Distance, chart = 'i')
qic(df10x$Calibrated_Distance, chart = 'i')
qic(df10x$MPC_Distance, chart = 'mr')
qic(df10x$Calibrated_Distance, chart = 'mr')

qic(df6x$Output_Change, x = df6x$QA_Date, chart = 'xbar')

qic(df6x$SymmInAT, chart = 'mr')

```


```{r}
qic(recent6x$SymmInAT, x = recent6x$QA_Date, chart = 'mr')
```
box cox transformation to normal distribution
```{r}
library(MASS)
help(boxcox)

boxcox(recent6x$SymmInAT~1)

```

