---
title: "MPC"
author: "NCompton"
date: "28/06/2021"
output: html_document
---

knit at end

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

install packages
```{r, echo = FALSE}
install.packages("odbc")
install.packages("DBI")
install.packages("dplyr")
install.packages("fitdistrplus")
install.packages("ggplot2")
install.packages("qicharts2")
install.packages("psych")
install.packages("distributions3")
install.packages("ggstatsplot")
install.packages("forecast")
install.packages("qcc")
install.packages("lubridate")
install.packages("Metrics")
```

Set libraries
```{r}
library(psych)
library(qicharts2)
library(ggplot2)
library(fitdistrplus)
library(dplyr)
library(odbc)
library(distributions3)
library(ggstatsplot)
library(forecast)
library(qcc)
library(lubridate)
library(Metrics)
```

Define functions

```{r}
# function to return 2-sample z statistic and p value
z.test <- function(x, y){
  # get the sample mean, sd and n
  mu_1 = mean(x)
  mu_2 = mean(y)
  sigma_1 = sd(x)^2
  sigma_2 = sd(y)^2
  n_1 = length(x)
  n_2 = length(y)
  
  # calculate 2 tailed z stat
  z_stat = (mu_1 - mu_2) / (sqrt ((sigma_1/n_1) + (sigma_2/n_2)))
  
  # calculate p value
  Z <- Normal(0,1) # make a standard normal r.v
  p = 2 * cdf(Z, abs(z_stat))
  return(c(z_stat, p))
  
}


# Paired Z tests
z.paired <- function(x, y, d0, alpha){
  if(missing(d0)){d0 <- 0}
  if(missing(alpha)){alpha <- 0.05}
  mean_diff <- mean(x-y)
  sigma_diff <- sqrt(var(x)+var(y)-2*cov(x,y))
  n <- length(x)
  z<-sqrt(n)*((mean_diff-d0)/sigma_diff)
  p_value=2*pnorm(-abs(z))
  
    # qnorm gives the alpha quantile of the standard normal  distribution
  # reject H0 that difference of means = d0 if z < z(alpha/2) or
  # z > z(1-alpha/2)
  
  
  if(z < qnorm(alpha/2)) {cat("Null hypothesis can be rejected. Evidence the difference in means is not equal to", d0)}
  if (z > qnorm(1 - (alpha/2))) {
  cat("Null hypothesis can be rejected. Evidence the difference in means is not equal to", d0)}
  if ( qnorm(alpha/2)< z & z < qnorm(1- (alpha/2))) {  cat("Null hypothesis can not be rejected, evidence the difference in means is equal to", d0)}
  
  return(c(z, p_value))
  
}


# Function to find the pearson's correlation between the variables
Symm_Flat_Rel <- function(df){
  print("Symmetry Crossline and Inline")
print(cor.test(df$SymmCrossAT, df$SymmInAT))

print("Uniformity and Inline Symmetry")
print(cor.test(df$Uniformity, df$SymmInAT))

print("Uniformity and Crossline Symmetry")
print(cor.test(df$Uniformity, df$SymmCrossAT))

print("Flatness Crossline and Inline")
print(cor.test(df$FlatInAT, df$FlatCrossAT))

print("Uniformity and Inline Flatness")
print(cor.test(df$Uniformity, df$FlatInAT))

print("Uniformity and Crossline Flatness")
print(cor.test(df$Uniformity, df$FlatCrossAT))

print("Output change and symmetry inline")
print(cor.test(df$Output_Change, df$SymmInAT))

print("Output change and symmetry crossline")
print(cor.test(df$Output_Change, df$SymmCrossAT))

print("Output change and flatness inline")
print(cor.test(df$Output_Change, df$FlatInAT))

print("Output change and flatness crossline")
print(cor.test(df$Output_Change, df$FlatCrossAT))

}


# Function to plot the relationships between variables
plot_cor <- function(df, energy){
  
  colours <- c("Inline" = "blue", "Crossline" = "yellow")
print(ggplot(df, aes(x = SymmInAT, y = SymmCrossAT)) +
  geom_point(color = "red") +
    geom_smooth(method='loess') +
    xlab("Symmetry Inline") + 
    ylab("Symmetry Crossline") + 
    labs(title = energy))

print(ggplot(df, aes(x = Uniformity)) +
  geom_point(aes(y = SymmInAT, color = "Inline")) +
  geom_point(aes(y = SymmCrossAT, color = "Crossline")) +
  geom_smooth(aes(x = Uniformity, y = SymmInAT, color = "Inline"), method = 'loess') +
  geom_smooth(aes(x = Uniformity, y = SymmCrossAT, color = "Crossline"), method = 'loess') +
  xlab("Uniformity") +
  ylab("Symmetry") +
    labs(title = energy, color = "legend") +
      scale_color_manual(values = colours))

print(ggplot(df, aes(Output_Change)) +
  geom_point(aes(y = SymmInAT, color = "Inline")) +
  geom_point(aes(y = SymmCrossAT, color = "Crossline")) +
     geom_smooth(aes(x = Output_Change, y = SymmInAT, color = "Inline"), method = 'loess') +
  geom_smooth(aes(x = Output_Change, y = SymmCrossAT, color = "Crossline"), method = 'loess') +
  xlab("Output Change") + 
  ylab("Symmetry") +
    labs(title = energy, color = "legend")+
    scale_color_manual(values = colours))

print(ggplot(df, aes(x = FlatInAT, y = FlatCrossAT)) +
  geom_point(color = "red") +
    geom_smooth(method='loess') +
   xlab("Flatness Inline") + 
  ylab("Flatness Crossline") +
    labs(title = energy))

print(ggplot(df, aes(Uniformity)) +
  geom_point(aes(y = FlatInAT, color = "Inline")) +
  geom_point(aes(y = FlatCrossAT, color = "Crossline")) +
     geom_smooth(aes(x = Uniformity, y = FlatInAT, color = "Inline"), method = 'loess') +
  geom_smooth(aes(x = Uniformity, y = FlatCrossAT, color = "Crossline"), method = 'loess') +
  xlab("Uniformity") + 
  ylab("Flatness") +
    labs(title = energy, color = "legend") +
    scale_color_manual(values = colours))

print(ggplot(df, aes(Output_Change)) +
  geom_point(aes(y = FlatInAT, color = "Inline")) +
  geom_point(aes(y = FlatCrossAT, color = "Crossline")) + 
  geom_smooth(aes(x = Output_Change, y = FlatInAT, color = "Inline"), method = 'loess') +
  geom_smooth(aes(x = Output_Change, y = FlatCrossAT, color = "Crossline"), method = 'loess') +
  xlab("Output Change") +
  ylab("Flatness") +
    labs(title = energy,  color = "legend") +
    scale_color_manual(values = colours))

print(ggplot(df, aes(x = MPC_Distance, y = Calibrated_Distance)) +
        geom_point(color = "red") +
        xlab("MPC Centre Shift") +
        ylab("Calibrated Centre Shift") +
        labs(title = energy))

}


# Function to calculate sensitivty and specificty
sens_spec <- function(var1, var2){

false_pos <- sum(var1 == "pass" & var2 == "fail")
true_pos <- sum(var1 == "pass" & var2 == "pass")
false_neg <- sum(var1 == "fail" & var2 == "pass")
true_neg <- sum(var1 == "fail" & var2 == "fail")

sens <- true_pos / (true_pos + false_neg)
cat("   Sensitivity is ", sens)

spec <- true_neg / (true_neg + false_pos)
cat("  Specificity is ", spec)
}


# Function to calculate senstivity and specficity for each variable
calc_sens_spec <- function(df){

  # symmetry against uniformity
  print(" Uniformity vs Combined Symmetry ")
  sens_spec(df$uniformity_test, df$symm_test)
  print(" Uniformity vs symmetry inline ")
  sens_spec(df$uniformity_test, df$symm_in_test)
  print("Uniformity vs symmetry crossline ")
  sens_spec(df$uniformity_test, df$symm_cross_test)
  
  # symmetry against beam output change
  print("Beam output vs symmetry")
  sens_spec(df$output_test, df$symm_test)
  print("Beam output vs symmetry inline")
  sens_spec(df$output_test, df$symm_in_test)
  print("Beam output vs symmetry crossline")
  sens_spec(df$output_test, df$symm_cross_test)
  
  # centre shift distance
  print("Beam centre shift: MPC vs calibrated")
  sens_spec(df$centre_mpc_test, df$centre_cal_test)
  
   # flatness against uniformity
  print(" Uniformity vs Combined Symmetry ")
  sens_spec(df$uniformity_test, df$flat_test)
  print(" Uniformity vs flatness inline ")
  sens_spec(df$uniformity_test, df$flat_in_test)
  print("Uniformity vs flatness crossline ")
  sens_spec(df$uniformity_test, df$flat_cross_test)
  
  # flatness against beam output change
  print("Beam output vs flatness")
  sens_spec(df$output_test, df$flat_test)
  print("Beam output vs flatness inline")
  sens_spec(df$output_test, df$flat_in_test)
  print("Beam output vs flatness crossline")
  sens_spec(df$output_test, df$flat_cross_test)
}



# Function to add pass/fail columns to the dataframe

add_cols <- function(df){
  
  symmetry_threshold <- 3
  flat_threshold <- 106
  distance_threshold <- 0.5
  
  # set the mpc thresholds
  centre_threshold <- 0.5
  uniformity_threshold <- 2
  output_threshold <- 2
  
  f <- `$<-`
  
  # uniformity
  df <- f(df, "uniformity_test", "fail")
  df$uniformity_test[df$Uniformity<uniformity_threshold & df$Uniformity>-uniformity_threshold] <- "pass"
  
    
  # combined symmetry
  # fail if 1 of them fails
  df<- f(df, "symm_test", "fail")
  df$symm_test[df$SymmInAT<symmetry_threshold & df$SymmInAT>-symmetry_threshold & df$SymmCrossAT<symmetry_threshold & df$SymmCrossAT>-symmetry_threshold] <- "pass"
  
   # fail for inline only
  df <- f(df, "symm_in_test", "fail")
  df$symm_in_test[df$SymmInAT<symmetry_threshold & df$SymmInAT>-symmetry_threshold] <- "pass"
  
    # fail for crossline only
  df <- f(df, "symm_cross_test", "fail")
  df$symm_cross_test[df$SymmCrossAT<symmetry_threshold & df$SymmCrossAT>-symmetry_threshold] <- "pass"
  
  
  
  # flatness
  # fail if 1 of them fails
  df<- f(df, "flat_test", "fail")
  df$flat_test[df$FlatInAT<flat_threshold & df$FlatInAT>-flat_threshold & df$FlatCrossAT<flat_threshold & df$FlatCrossAT>-flat_threshold] <- "pass"
  
   # fail for inline only
  df <- f(df, "flat_in_test", "fail")
  df$flat_in_test[df$FlatInAT<flat_threshold & df$FlatInAT>-flat_threshold] <- "pass"
  
    # fail for crossline only
  df <- f(df, "flat_cross_test", "fail")
  df$flat_cross_test[df$FlatCrossAT<flat_threshold & df$FlatCrossAT>-flat_threshold] <- "pass"
  
  
    # output change
  df <- f(df, "output_test", "fail")
  df$output_test[df$Output_Change<output_threshold & df$output_test>-output_threshold] <- "pass"
  
  # centre shift
  df <- f(df, "centre_mpc_test", "fail")
  df$centre_mpc_test[df$MPC_Distance<distance_threshold & df$MPC_Distance>-distance_threshold] <- "pass"
  
  df <- f(df, "centre_cal_test", "fail")
  df$centre_cal_test[df$Calibrated_Distance<centre_threshold & df$Calibrated_Distance>-centre_threshold] <- "pass"
  
  return(df)
  
}


```


max difference function
```{r}

difference <- function(df){
  
  # uniformity
cat(" Max difference symmetry inline and uniformity is ", max( abs(df$SymmInAT - df$Uniformity)))
  
  cat(" Max differencesymmetry crossline and uniformity is ", max( abs(df$SymmCrossAT - df$Uniformity)))
  
  cat(" Max difference flatneess inline and uniformity is ", max( abs( (df$FlatInAT-100) - df$Uniformity)))
  
  cat(" Max difference flatness crossline and uniformity is ", max( abs( (df$FlatCrossAT-100) - df$Uniformity)))
  
  
  
  # output change
  cat(" Max difference symmetry inline and output change is ", max( abs(df$SymmInAT - df$Output_Change)))
  
  cat(" Max difference symmetry crossline and output change is ", max( abs(df$SymmCrossAT - df$Output_Change)))
  
  cat(" Max difference flatness inline and output change is ", max( abs( (df$FlatInAT-100) - df$Output_Change)))
  
  cat(" Max difference flatness crossline and output change is ", max( abs( (df$FlatCrossAT-100) - df$Output_Change)))
  
  
  
  # beam centre shift
  cat(" Max difference beam centre shift is ", max( abs(df$MPC_Distance - df$Calibrated_Distance)))

  
}


```


Set up the database connection
```{r}
con <- DBI::dbConnect(odbc::odbc(), 
                      Driver = "SQL Server", 
                      Server = "IT049561\\SQLEXPRESS", 
                      Database = "MPC", 
                      Trusted_Connection = "True")
```


Get a dataframe for each beam energy
```{sql connection=con, output.var = "df6x"}
SELECT *
FROM "6x"
WHERE QA_Date >= '2018-01-01'
```
```{sql connection=con, output.var = "df10x"}
SELECT *
FROM "10x"
WHERE QA_Date >= '2018-01-01'

```
```{sql connection=con, output.var = "df10fff"}
SELECT *
FROM "10fff"
WHERE QA_Date >= '2018-01-01'

```

Get the dataframes of the recent data - from 1st April 2021 when we measured the water phantom
```{sql connection=con, output.var=recent6x}
SELECt *
FROM "6x"
WHERE QA_Date >= '2021-04-01'
```

```{sql connection=con, output.var=recent10x}
SELECt *
FROM "10x"
WHERE QA_Date >= '2021-04-01'
```

```{sql connection=con, output.var=recent10fff}
SELECt *
FROM "10fff"
WHERE QA_Date >= '2021-04-01'
```



Remove outliers from the recent dfs so that we don't skew the results? Particularly for 10FFF. While these data points are valid, they are going to mess up all the results - we know the flattening free will have bigger fluctuations but retaining them bias's the analysis; it removes the actual relationship from sight - SD and mean are sensitive to outliers and look at the boxplot - we can't see most of the spread - including outliers means that later we can't see the correlations.

Remove outliers - visual inspection to check the outliers have been removed successfully/

```{r}


remove_outliers <- function(df, cols = names(df)) {
  for (col in cols) {
    boxplot(df[[col]], plot=FALSE)$out
    outliers <- boxplot(df[[col]], plot=FALSE)$out
    df <- df[-which(df[[col]] %in% outliers),]
  }
  df
}

# we want to remove outliers from all 4 flatness columns of 10FFF beam

rm_out_10fff <- remove_outliers(recent10fff, c("FlatInAT", "FlatCrossAT", "FlatInBT", "FlatCrossBT"))

```



Descriptive statistics with psych
```{r, echo = FALSE}
describeBy(df6x$MPC_Distance)

describeBy(df6x$Calibrated_Distance)

describeBy(df6x$Uniformity)

describeBy(df6x$Output_Change)
```


Define the random samples
```{r}
set.seed(1) # set seed makes the analysis reproducible
rand_6x <- recent6x[sample(nrow(recent6x), size=50), ]
rand_10x <- recent10x[sample(nrow(recent10x), size=50), ]
rand_10fff <- recent10fff[sample(nrow(recent10fff), size=50), ]
rand_10fff_rm_out <- rm_out_10fff[sample(nrow(rm_out_10fff), size=50),]

```


Z tests 
Symmetry inline and crossline directions after calibration

Z test null hypothesis: the means of the two groups are equal
The Z test statistic is inside the 95% confidence interval, so we can not reject the null hypothesis. 
The p-value tells us how statistically significant this is and it is quite large; but still we don't have enough evidence to reject the null hypothesis.



Comparing Symmetry and Flatness in each direction
```{r}

boxplot(recent6x$SymmInAT, recent6x$SymmCrossAT, recent10x$SymmInAT, recent10x$SymmCrossAT, recent10fff$SymmInAT, recent10fff$SymmCrossAT, main = "Symmetry (%)", names = c("6x", "6x", "10x", "10x", "10fff", "10fff"), col = c(2,3,2,3,2,3))
legend(x = "topright", legend = c("Inline", "Crossline"), fill = 2:3, col = c("orange", "red"))


boxplot(recent6x$FlatInAT, recent6x$FlatCrossAT, recent10x$FlatInAT, recent10x$FlatCrossAT, main = "Flatness", names = c("6x", "6x", "10x", "10x"), col = c(2,3,2,3))
legend(x = "topright", legend = c("Inline", "Crossline"), fill = 2:3, col = c("orange", "red"))

boxplot(rm_out_10fff$FlatInAT, rm_out_10fff$FlatCrossAT, main = "Flatness", names = c("10xFFF", "10xFFF"), col = c(2,3))
legend(x = "topright", legend = c("Inline", "Crossline"), fill = 2:3, col = c("orange", "red"))

# Compare symm and flat
# 6x
z.paired(rand_6x$SymmInAT, rand_6x$SymmCrossAT, 0, alpha = 0.05)
z.paired(rand_6x$FlatInAT, rand_6x$FlatCrossAT, 0, alpha = 0.05)

# 10x
z.paired(rand_10x$SymmInAT, rand_10x$SymmCrossAT, 0, alpha = 0.05)
z.paired(rand_10x$FlatInAT, rand_10x$FlatCrossAT, 0, alpha = 0.05)

# 10xFFF
z.paired(rand_10fff_rm_out$SymmInAT, rand_10fff_rm_out$SymmCrossAT, 0, alpha = 0.05)
z.paired(rand_10fff_rm_out$FlatInAT, rand_10fff_rm_out$FlatCrossAT, 0, alpha = 0.05)
```


Paired Z tests to see if there's a significant difference as a result of applying the calibration transformation.

Symmetry
```{r}
boxplot(recent6x$SymmInBT, recent6x$SymmInAT, recent10x$SymmInBT, recent10x$SymmInAT, recent10fff$SymmInBT, recent10fff$SymmInAT, main = "Symmetry Inline (%)", names = c("6x", "6x", "10x", "10x", "10fff", "10fff"), col = c(2,3,2,3,2,3))
legend(x = "topright", legend = c("Before Calibration", "After Calibration"), fill = 2:3, col = c("orange", "red"))

boxplot(recent6x$SymmCrossBT, recent6x$SymmCrossAT, recent10x$SymmCrossBT, recent10x$SymmCrossAT, recent10fff$SymmCrossBT, recent10fff$SymmCrossAT, main = "Symmetry Crossline (%)", names = c("6x", "6x", "10x", "10x", "10fff", "10fff"), col = c(2,3,2,3,2,3))
legend(x = "topright", legend = c("Before Calibration", "After Calibration"), fill = 2:3, col = c("orange", "red"))

# 6x
z.paired(rand_6x$SymmInBT, rand_6x$SymmInAT, 0, alpha = 0.05)
z.paired(rand_6x$SymmCrossBT, rand_6x$SymmCrossAT, 0, alpha = 0.05)

# 10x
z.paired(rand_10x$SymmInBT, rand_10x$SymmInAT, 0, alpha = 0.05)
z.paired(rand_10x$SymmCrossBT, rand_10x$SymmCrossAT, 0, alpha = 0.05)

# 10fff
z.paired(rand_10fff$SymmInBT, rand_10fff$SymmInAT, 0, alpha = 0.05)
z.paired(rand_10fff$SymmCrossBT, rand_10fff$SymmCrossAT, 0, alpha = 0.05)

```

Flatness
```{r}

boxplot(recent6x$FlatInBT, recent6x$FlatInAT, recent10x$FlatInBT, recent10x$FlatInAT, main = "Flatness Inline", names = c("6x", "6x", "10x", "10x"), col = c(2,3,2,3))
legend(x = "topright", legend = c("Before Calibration", "After Calibration"), fill = 2:3, col = c("orange", "red"))

boxplot(rm_out_10fff$FlatInBT, rm_out_10fff$FlatInAT, rm_out_10fff$FlatCrossBT, rm_out_10fff$FlatCrossAT, main = "Flatness: 10xFFF", names = c("Inline", "Inline", "Crossline", "Crossline"), col = c(2,3,2,3))
legend(x = "topleft", legend = c("Before", "After"), fill = 2:3, col = c("orange", "red"))

boxplot(recent6x$FlatCrossBT, recent6x$FlatCrossAT, recent10x$FlatCrossBT, recent10x$FlatCrossAT, main = "Flatness Crossline", names = c("6x", "6x", "10x", "10x"), col = c(2,3,2,3))
legend(x = "topright", legend = c("Before Calibration", "After Calibration"), fill = 2:3, col = c("orange", "red"))

# 6x
z.paired(rand_6x$FlatInBT, rand_6x$FlatInAT, 0, alpha = 0.05)
z.paired(rand_6x$FlatCrossBT, rand_6x$FlatCrossAT, 0, alpha = 0.05)

# 10x
z.paired(rand_10x$FlatInBT, rand_10x$FlatInAT, 0, alpha = 0.05)
z.paired(rand_10x$FlatCrossBT, rand_10x$FlatCrossAT, 0, alpha = 0.05)

# 10fff
#z.paired(rand_10fff$FlatInBT, rand_10fff$FlatInAT, 0, alpha = 0.05)
#z.paired(rand_10fff$FlatCrossBT, rand_10fff$FlatCrossAT, 0, alpha = 0.05)

# 10fff with outliers removed, random sample
z.paired(rand_10fff_rm_out$FlatInBT, rand_10fff_rm_out$FlatInAT)
z.paired(rand_10fff_rm_out$FlatCrossBT, rand_10fff_rm_out$FlatCrossAT)
```


Compare the beam centre shift
```{r}

boxplot(recent6x$MPC_Distance, recent6x$Calibrated_Distance, recent10x$MPC_Distance, recent10x$Calibrated_Distance, recent10fff$MPC_Distance, recent10fff$Calibrated_Distance, main = "Beam Centre Shift (mm)", names = c("6x", "6x", "10x", "10x", "10fff", "10fff"), col = c(2,3,2,3,2,3))
legend(x = "topleft", legend = c("MPC Distance", "Calibrated Distance"), fill = 2:3, col = c("orange", "red"))


z.paired(rand_6x$MPC_Distance, rand_6x$Calibrated_Distance, 0.5, alpha = 0.05)

z.paired(rand_10x$MPC_Distance, rand_10x$Calibrated_Distance, 0.5, alpha = 0.05)

z.paired(rand_10fff$MPC_Distance, rand_10fff$Calibrated_Distance, 0.5, alpha = 0.05)

```


Sensitivity and Specificity


add pass/fail columns to the dataframe
```{r}
recent6x <- add_cols(recent6x)
recent10x <- add_cols(recent10x)
recent10fff <- add_cols(recent10fff)

```

6x 

```{r}
calc_sens_spec(recent6x)
```
10x

```{r}
calc_sens_spec(recent10x)

```

```{r}
calc_sens_spec(recent10fff)
```


Difference between symmetry and uniformity
Compare how much difference there is overall - e.g. all symmetry values are within x% of uniformity


6x
```{r}
difference(recent6x)
```

10x
````{r}
difference(recent10x)
```

10fff
```{r}
difference(recent10fff)
```


Are there a correlations between symmetry, flatness and uniformity, output change?
With time?

Pearson's test for correlation

```{r}
Symm_Flat_Rel(recent6x)
```
```{r}
Symm_Flat_Rel(recent10x)
```
```{r}
Symm_Flat_Rel(recent10fff)
```


```{r}
plot_cor(recent6x, energy = "6x")
plot_cor(recent10x, energy = "10x")
plot_cor(recent10fff, energy = "10xFFF")
plot_cor(rm_out_10fff, energy = "10xFFF (outliers removed)")

```

Time series

Plot the flatness and symmetry against time
```{r}
plot_time <- function(df, energy){
  colours <- c("Inline" = "blue", "Crossline" = "yellow")
  
print(ggplot(df, aes(x = QA_Date)) +
  geom_point(aes(y = FlatInAT, color = "Inline")) + 
  geom_point(aes(y = FlatCrossAT, color = "Crossline")) +
  geom_smooth(aes(x = QA_Date, y = FlatInAT, color = "Inline"), method = 'loess') +
   geom_smooth(aes(x = QA_Date, y = FlatCrossAT, color = "Crossline"), method = 'loess') +
  labs(title = paste("Flatness", energy)))


print(ggplot(df, aes(x = QA_Date)) +
  geom_point(aes(y = SymmInAT, color = "Inline")) +
  geom_point(aes(y = SymmCrossAT, color = "Crossline")) +
    geom_smooth(aes(x = QA_Date, y = SymmInAT, color = "Inline"), method = 'loess') +
   geom_smooth(aes(x = QA_Date, y = SymmCrossAT, color = "Crossline"), method = 'loess') +
  labs(title = paste("Symmetry", energy)))
}

plot_time(recent6x, "6x")
plot_time(recent10x, "10x")
plot_time(recent10fff, "10xFFF")
plot_time(rm_out_10fff, "10xFFF (outliers removed)")

```

MPC results time series

Plot the beam output change, uniformity and centre shift over time
```{r}

plot_mpc_time <- function(df, energy){
  
  print(ggplot(df, aes(QA_Date, Output_Change)) +
  geom_point() +
    #geom_smooth(aes(x = QA_Date, y = Output_Change), method = "gam") +
  xlab("QA Date") +
  ylab("Output Change") +
  ggtitle(paste("Output Changes against Time", energy)))
  
    print(ggplot(df, aes(QA_Date, Uniformity)) +
  geom_point() +
    #geom_smooth(aes(x = QA_Date, y = Uniformity), method = "gam") +
  xlab("QA Date") +
  ylab("Uniformity") +
  ggtitle(paste("Uniformity against Time", energy)))
    
    print(ggplot(df, aes(QA_Date, MPC_Distance)) +
  geom_point() +
    #geom_smooth(aes(x = QA_Date, y = MPC_Distance), method = "gam") +
  xlab("QA Date") +
  ylab("Beam Centre Shift (mm)") +
  ggtitle(paste("Centre shift against Time", energy)))


}

```

MPC results against time
```{r}

#plot_mpc_time(df6x, "6x")
#plot_mpc_time(df10x, "10x")
#plot_mpc_time(df10fff, "10xFFF")


# clean the df to remove the outliers identified visually from the graphs above
df10x <- df10x[which(df10x$Output_Change > -5),]
df10x <- df10x[which(df10x$Uniformity < 10),]
df10x <- df10x[which(df10x$MPC_Distance < 0.5),]


df10fff <- df10fff[which(df10fff$Output_Change<9),]
df10fff <- df10fff[which(df10fff$Output_Change> -5),]
df10fff <- df10fff[which(df10fff$Uniformity<40),]
df10fff <- df10fff[which(df10fff$MPC_Distance<0.5),]


plot_mpc_time(df6x, "6x")
plot_mpc_time(df10x, "10x")
plot_mpc_time(df10fff, "10xFFF")
```
centre shift of MPC and my algorithm - shows bias
```{r}
plot_centre_shift <- function(df, energy){
  colors <- c("MPC" = "blue", "Calibrated" = "red")
  
    print(ggplot(df, aes(x = QA_Date)) +
  geom_point(aes(y = MPC_Distance, color = "MPC")) +
    geom_point(aes(y = Calibrated_Distance, color = "Calibrated")) +
    geom_smooth(aes(x = QA_Date, y = MPC_Distance, color = "MPC"), method = "loess") +
    geom_smooth(aes(x = QA_Date, y = Calibrated_Distance, color = "Calibrated"), method = "loess") +
  xlab("QA Date") +
  ylab("Beam Centre Shift (mm)") +
  ggtitle(paste("Centre shift against Time", energy)))
}

plot_centre_shift(recent6x, "6x")
plot_centre_shift(recent10x, "10x")
plot_centre_shift(recent10fff, "10xFFF")

```


Predicting future values

Set up the training and testing dfs
Set up the UCL, LCL, CL
```{r}

training_df <- function(df){
  # first 85% of the data is training
  return(head(df, round(nrow(df) * 0.85)))
}

testing_df <- function(df){
  # last 15% of the data is testing
  return(tail(df, nrow(df) - round(nrow(df) * 0.85)))
}

training_df6x <- training_df(df6x)
testing_df6x <- testing_df(df6x)


# control limits
limits <- function(dfcol){

  the_Mean <- mean(dfcol)
  the_mR <- mean(abs(diff(dfcol)))
  the_Sigma <- the_mR/1.128
  the_LCL <- the_Mean - 3 * the_Sigma
  the_UCL <- the_Mean + 3 * the_Sigma
  
  return(c(the_Mean, the_LCL, the_UCL))
}


```

```{r}
av_accuracy <- function(df, state){
  
  if(state == "fail"){ 
    non_state_1 <- "pass"
    non_state_2 <- "warning"}
  
   if(state == "warning"){ 
     non_state_1 <- "pass"
     non_state_2 <- "fail"}
  
   if(state == "pass"){ 
     non_state_1 <- "warning"
     non_state_2 <- "fail"}
  
  true_pos <-  sum(df$pred_state == state & df$actual_state == state)
  
  true_neg <- sum(df$pred_state == non_state_1 & df$actual_state == non_state_1) + sum(df$pred_state == non_state_2 & df$actual_state == non_state_2)
  
  false_pos <- sum(df$pred_state == state & df$actual_state != state)
  
  false_neg <- sum(df$pred_state == non_state_1 & df$actual_state != non_state_1) + sum(df$pred_state == non_state_2 & df$actual_state != non_state_2)
  
  acc <- (true_pos + true_neg) / (true_pos+true_neg+false_pos+false_neg)
  
  av_acc <- acc / nrow(df)
  return(av_acc)
}
```


Forecasting with ARIMA

convert this into a function 
repeat analysis with other mpc metrics

" The beam output change has high variability over time and is recorded as a nonstationary time-series dataset, which affects the accuracy of the predictive model. Resolution of this problem may require a different predictive model. Examples of possible models include the EWMA model, recurrent neural networks,21 and artificial neural networks,22 which could be investigated in future work."

from the paper using ARIMA and SPC
so should I try to implement one of those?

```{r}
# set up a vector which calcs the limits iteratively for the actual uniformity values

forecating_df <- function(trainingdf, testingdf, datecol, threshold){
  
cl <- c()
lcl <- c()
ucl <- c()

actual <- c(trainingdf)
predicted <- c()
trend <- c()

# iteratively construct the dataframe from predicted values
for (n in 1:length(testingdf)){
  # add the actual value
  actual <- c(actual, testingdf[n])
  
  cl <- c(cl, limits(actual)[1])
  lcl <- c(lcl, limits(actual)[2])
  ucl <- c(ucl, limits(actual)[3])
  
  fit <- auto.arima(actual)
  pred <- forecast(fit, 1)
  
  if ( (n-1) %% 5 == 0){
  pred_5 <- forecast(fit, 5)
  trend <- c(trend, pred_5$mean)
  }
  
  predicted <- c(predicted, pred$mean)
  
}
  
  
  arima_df <- data.frame(actual_val <- c(testingdf),
                        central <- cl,
                        lower <- lcl,
                        upper <- ucl,
                        one_step <- predicted,
                        five_step <- trend[1:length(testingdf)],
                        Date <- c(datecol))
  
  
  # plot values
  print(ggplot(arima_df, aes(x = Date)) +
  geom_point(aes(y = actual_val), color = "blue") +
  geom_point(aes(y = one_step), color = "red") +
  geom_point(aes(y = lcl), color = "yellow") +
  geom_point(aes(y = ucl), color = "yellow") +
  geom_point(aes(y = five_step), color = "green"))
  
  # calculate errors
  # root mean squared error comparing the actual to the predicted
  error <- rmse(arima_df$actual_val, arima_df$one_step)
  abs_error <- mae(arima_df$actual_val, arima_df$one_step)
  cat("Root mean squared error is ", error)
  cat("  Absolute error is ", abs_error)
  
  
  # add a state pass/warning/fail to the df
  f <- `$<-`
  
  # uniformity
  arima_df <- f(arima_df, "pred_state", "warning")
  arima_df$pred_state[arima_df$one_step <= -threshold] <- "fail"
  arima_df$pred_state[arima_df$one_step>threshold] <- "fail"
  arima_df$pred_state[arima_df$one_step<arima_df$upper & arima_df$one_step>arima_df$lower] <- "pass"
  
  arima_df <- f(arima_df, "actual_state", "warning")
  arima_df$actual_state[arima_df$actual_val <= -threshold] <- "fail"
  arima_df$actual_state[arima_df$actual_val>threshold] <- "fail"
  arima_df$actual_state[arima_df$actual_val<arima_df$upper & arima_df$actual_val>arima_df$lower] <- "pass"
  
  # calculate average accuracy
  average_accuracy <- av_accuracy(arima_df, "warning")
  cat("  Average accuracy is ", average_accuracy)
  
  return(arima_df)

}
  


```

```{r}
uniformdf <- forecating_df(training_df6x$Uniformity, testing_df6x$Uniformity, testing_df6x$QA_Date, 2)
```






